Lightness constancy under fixed geometry

### 2018-01-04 Draft1/draft1.tex
Today’s work

A difficulty in using machine learning approaches for color constancy is that these approaches require large training sets with images of scenes containing objects with known surface reflectances illuminated with known illuminations.  Such data sets are not readily available. Although there are several databases of calibrated color natural images, these do not generally provide the corresponding information about surface and illuminant ground truth \cite{ChakrabartiHyperspectral,NascimentoFoster2016,ParragaHyperspectralData,TkacikUpennHypersepctralData,skauli2013collection},  ( ADD McGill Database). Generating databases with both the images and the ground truth scene information is not trivial. One approach has been to acquire images where objects of known reflectance have been inserted into the scene, and to use the image of these reference objects to provide an estimate of the illumination impinging at various scene locations \cite{NascimentoFoster2016}.  This is somewhat painstaking, however, and in any case provides information about the illumination only at a small number of selected locations, so that generalizing across the image requires strong assumptions. Some progress is being made on characterizing how illumination varies over space in natural scenes \cite{mury2007spatial}, which in the longer run may guide how information from reference objects can be used to provide reasonable estimates.  Another approach is to acquire images of posed scenes, where the surface reflectances of objects are measured individually for many of the objects in the scene \cite{funt1988color,ciurea2003large} (ADD UCSB hyperspectral). These data sets have been useful for evaluating computational color constancy algorithms, but the use of posed scenes does not scale well for use with supervised learning methods which needs large training sets.

@@

Because of the difficulty in obtaining large labelled data sets of natural images for use in developing and evaluating computational color constancy algorithms, in this paper we adopt the approach of using high-quality computer graphics to render images of what we refer to as {\em virtual natural scenes}. Here, in addition to employing computer graphics to render the light reaching the eye from specified scenes, we also model the initial steps of the human visual system, so that our supervised learning methods are applied not to image pixels but rather to modeled responses of the foveal cone photoreceptor mosaic. We describe this approach in detail below, but the key advantage is that we can produce large numbers of naturalistic images and the corresponding cone mosaic responses, where all of the underlying scene properties are well-specified. This allows us to explore and evaluate computational color constancy with rich stimuli, while retaining the ability to precisely control and manipulate the intrinsic properties of the objects and illuminants that comprise the scene. This general approach has been adopted by other labs for the study of computational estimation of optical flow \cite{baker2011database} and is becoming increasingly popular (OTHER REFS: Cornell graphics person, Gallant, Yamins, Autonomous vehicle work), as well as for the generation of stimuli for the study of human perception \cite{boyaci2003effect}(REFS: Many -- go to Boynton lecture slides for some).  Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the \nameref{Discussion}.

@@

In this work, we are interested in how well lightness constancy may be achieved using reasonable information that is local to the image region containing the target patch. 
% MAYBE SAY A LITTLE MORE ABOUT WHY - THIS IS A PRINCIPLED STANCE TAKEN IN MOST OF JOHANNES WORK.  OR MAYBE DEFER THAT TO THE DISCUSSION.
To this end, we crop a 51x51 pixel portion of the full rendered multispectral image that contains the target object and use this for further analysis (\ref{fig:croppedImage}).
Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this paper.

Next, the cropped multispectral images are used to simulate retinal cone responses using \href{https://github.com/isetbio}{Isetbio} software. The key steps we model are optical blurring, including axial chromatic aberration, in the formation of the retinal image, spatial sampling by an interleaved mosaic of L, M and S cones, and the Poisson nature of photopigment isomerization.  
% NEED TO WRITE IN THE KEY PARAMETERS USED. PROBABLY MARIMONT AND WANDELL OPTICS, STOCKMAN-SHARPE LMS CONE SPECTRAL SENSITIVITIES, THE L, M, S CONE RATIO IN THE MOSAIC, RECTANGULAR AND REGULAR CONE PACKING, SOME IMPLIED SPATIAL SCALE, AND SOME OVERALL SCALING THAT DRIVES THE MEAN LUMINANCE OF THE RETINAL IMAGES AND THUS THE AMOUNT OF POISSON NOISE.

% WE GOT TO HERE ON DECEMBER 21.
To generate the retinal response one needs to specify the retinal cone mosaic and the size of the retinal image. The retinal cone mosaic is specified in terms of the total number of cones and the proportion of the long(L), medium(M) and the short(S) cones in the mosaic. With these parameters, the software generates the specified number of LMS cones with random positions on a square grid. The size of the retinal image is specified in terms of the field of view and object distance from the eye.

We simulate visual system response using \href{http://isetbio.org}{Isetbio} software toolbox. Isetbio simulates a model of the visual system accounting for cornea, pupil size, lens and physiological optics. We use the multispectral cropped image (Fig.~\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with 51x51 cones in the l:m:s ratio 0.6:0.3:0.1. The response of the cones are measured as the number of isomerizations in the cone segment ({\it over what time ??}). The cone response functions are normalized by the total area under the response curves to model the amplification of the S cone response downstream and thus have comparable number if isomerizations in L, M and S cones. Further, to get the response of each cone class at each position on the retinal mosaic, we interpolate the response of each cone class. Thus we get a $51 \times 51$ cone response image for each cone class.

The three $51 \times 51$ cone response matrix corresponding to L, M and S cones are then reshaped into a $7803 \times 1$ vector. To model the contrast adaptation in the visual system, we perform a contrast normalization of the vector by subtracting and dividing by the mean. We further  rescale the vector to have norm 1. The cone-response vectors corresponding to the N images for each case are concatenated into a column-wise contrast normalized matrix $C_{7803\times N}$. The corresponding labels are represented by the row vector $Y_{1\times N}$.

@@

\subsubsection*{Linear Regression} For linear regression, we solve the equation $A_{1\times3}*C_{3\times N} = Y_{1\times N}$ for the regression parameters $A_{1\times3}$. Here, $C_{3\times N}$ is a matrix containing the contrast normalized L, M and S cone responses of the central patch of each image. The columns corresponds to the N images in the dataset. The rows correspond to the contrast normalized L, M and S cones responses averaged over a 3$\times$3 pixel central region of the image containing the target object.  The (unknown) regression parameters $A$ are obtained using Matlab linear regression function.

\subsubsection*{SVD regression}
SVD regression is a generalization of principal component analysis for labeled data \cite{ranganNotes}. Briefly, given an image dataset matrix $C_{M \times N}$ and labels $Y_{1\times N}$, the method finds unit vectors such that the projection of the data along this vector maximizes the separation of the images with different labels, while clustering the images with similar labels together. Mathematically, the method searches for a direction $\hat{u}$ that minimize the cost function:
\begin{align}
S = \sum\limits_{j,j'}\left(\hat{u}^\intercal C(:,j) - \hat{u}^\intercal C(:,j')\right)^2 \left(Y(j)-Y(j') \right)^2,
\end{align} 
where $C(:,j)$ represents the $j$'th column/image of the data matrix $C$. The direction of minimum cost is given by the eigenvector corresponding to the largest eigenvalue of the following matrix:
\begin{align}
N C Y_D^2 C^\intercal - 2(Y*\mathbbm{1})C Y_D C^\intercal + (Y*Y^\intercal) (CC^\intercal)-C Y_{D}^2\mathbbm{1}\mathbbm{1}^\intercal C^\intercal +2 C Y_{D} \mathbbm{1}\mathbbm{1}^\intercal Y_{D} C^\intercal - C \mathbbm{1}\mathbbm{1}^\intercal Y_{D}^2 C^\intercal.
\end{align}
Here, $Y_{D}$ is a diagonal matrix with the elements of the labels vector $Y$ along the diagonal, and $\mathbbm{1}$ is a $N\times 1$ vector with each element equal to one.

\subsubsection*{Accuracy maximization analysis}
AMA \cite{geisler2009optimal,burge2017accuracy} is a Bayesian ideal observer model for task specific dimensionality reduction. Given a labeled training set, a noisy response model, and a cost function, AMA returns a set of linear receptive fields that are optimal for performing a task. The task is implicitly defined in terms of the labels. The optimal filters that are returned by AMA are filters whose noisy response to the stimuli, as given by the specified response model, minimizes the cost function over the training set.

>>>>

A difficulty in using machine learning approaches for color constancy is that these approaches require large training sets with images of scenes containing objects with known surface reflectances illuminated with known illuminations.  Such data sets are not readily available. Although there are several databases of calibrated color natural images, these do not generally provide the corresponding information about surface and illuminant ground truth \cite{ChakrabartiHyperspectral,NascimentoFoster2016,ParragaHyperspectralData,TkacikUpennHypersepctralData,skauli2013collection}.
%http://tabby.vision.mcgill.ca, Olmos, A., Kingdom, F. A. A. (2004). A biologically inspired algorithm for the recovery of shading and reflectance images, Perception, 33, 1463 - 1473.
 Generating databases with both the images and the ground truth scene information is not trivial. One approach has been to acquire images where objects of known reflectance have been inserted into the scene, and to use the image of these reference objects to provide an estimate of the illumination impinging at various scene locations \cite{NascimentoFoster2016}.  This is somewhat painstaking, however, and in any case provides information about the illumination only at a small number of selected locations, so that generalizing across the image requires strong assumptions. Some progress is being made on characterizing how illumination varies over space in natural scenes \cite{mury2007spatial}, which in the longer run may guide how information from reference objects can be used to provide reasonable estimates.  Another approach is to acquire images of posed scenes, where the surface reflectances of objects are measured individually for many of the objects in the scene \cite{funt1988color,ciurea2003large}
%http://color.psych.upenn.edu/hyperspectral/
These data sets have been useful for evaluating computational color constancy algorithms, but the use of posed scenes does not scale well for use with supervised learning methods which needs large training sets.

@@

Because of the difficulty in obtaining large labelled data sets of natural images for use in developing and evaluating computational color constancy algorithms, in this paper we adopt the approach of using high-quality computer graphics to render images of what we refer to as {\em virtual natural scenes}. Here, in addition to employing computer graphics to render the light reaching the eye from specified scenes, we also model the initial steps of the human visual system, so that our supervised learning methods are applied not to image pixels but rather to modeled responses of the foveal cone photoreceptor mosaic. We describe this approach in detail below, but the key advantage is that we can produce large numbers of naturalistic images and the corresponding cone mosaic responses, where all of the underlying scene properties are well-specified. This allows us to explore and evaluate computational color constancy with rich stimuli, while retaining the ability to precisely control and manipulate the intrinsic properties of the objects and illuminants that comprise the scene. This general approach has been adopted by other labs for the study of computational estimation of optical flow \cite{baker2011database} and is becoming increasingly popular (OTHER REFS),
% Kavita Bale at Cornell, Check her website
% Autonmous vehicle work - ask BAW
as well as for the generation of stimuli for the study of human perception \cite{boyaci2003effect}(REFS: Many -- go to Boynton lecture slides for some).
% Add "e.g." in front of citation list, however you do that in BibTex
%@article{todd1983perception,
%  title={Perception of surface curvature and direction of illumination from patterns of shading.},
%  author={Todd, James T and Mingolla, Ennio},
%  journal={Journal of Experimental Psychology: Human perception and performance},
%  volume={9},
%  number={4},
%  pages={583},
%  year={1983},
%  publisher={American Psychological Association}
%}
%@article{RN456,
%   author = {Arend, L and Reeves, A},
%   title = {Simultaneous color constancy},
%   journal = {Journal Of The Optical Society Of America A},
%   volume = {3},
%   number = {10},
%   pages = {1743-1751},
%   DOI = {10.1364/josaa.3.001743},
%   url = {http://josaa.osa.org/abstract.cfm?URI=josaa-3-10-1743},
%   year = {1986},
%   type = {Journal Article}
%}
%@article{johnston1993integration,
%  title={Integration of depth modules: Stereopsis and texture},
%  author={Johnston, Elizabeth B and Cumming, Bruce G and Parker, Andrew J},
%  journal={Vision research},
%  volume={33},
%  number={5},
%  pages={813--826},
%  year={1993},
%  publisher={Elsevier}
%}
Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the \nameref{Discussion}.

@@

In this work, we are interested in how well lightness constancy may be achieved using reasonable information that is local to the image region containing the target patch. 
% MAYBE SAY A LITTLE MORE ABOUT WHY - THIS IS A PRINCIPLED STANCE TAKEN IN MOST OF JOHANNES WORK.  OR MAYBE DEFER THAT TO THE DISCUSSION.
To this end, we crop a 51x51 pixel portion of the full rendered multispectral image that contains the target object and use this for further analysis (\ref{fig:croppedImage}).
Fig. \ref{fig:studiedCases} shows examples of such cropped image at 3 lightness levels for the three cases considered in this paper.
Next, the cropped multispectral images are used to simulate retinal cone responses using \href{https://github.com/isetbio}{Isetbio} software. The key steps we model are the size of the pupil, optical blurring (including axial chromatic aberration) in the formation of the retinal image, spatial sampling by an interleaved mosaic of long (L), middle (M)  and short (S) wavelength-sensitive cones, and the Poisson nature of photopigment isomerization.
We use the multispectral cropped image (Fig.~\ref{fig:croppedImage}) as the optical stimulus to simulate the response of the cones in the retina. The cropped image patch is assumed to cover a $1^{\circ}$ field of view at a distance of 1 m. The retinal cone mosaic is modeled as a square grid with $51 \times 51$ cones in the L:M:S ratio 0.6:0.3:0.1. The response of the cones are taken as the number of isomerizations in the cone segment ({\it over what time ??}).

% NEED TO WRITE IN THE KEY PARAMETERS USED. PROBABLY MARIMONT AND WANDELL OPTICS, STOCKMAN-SHARPE LMS CONE SPECTRAL SENSITIVITIES, THE L, M, S CONE RATIO IN THE MOSAIC, RECTANGULAR AND REGULAR CONE PACKING, SOME IMPLIED SPATIAL SCALE, AND SOME OVERALL SCALING THAT DRIVES THE MEAN LUMINANCE OF THE RETINAL IMAGES AND THUS THE AMOUNT OF POISSON NOISE. ALSO INTEGRATION TIME.
The cone response functions are normalized by the total area under the response curves to model the amplification of the S cone response downstream and thus have comparable number if isomerizations in L, M and S cones. 
% NEED TO UNDERSTAND THIS.  WHEN WAS THE GAIN, BEFORE OR AFTER NOISE ADDED, AND LET'S REMEMBER WHY WE DID IT.

The isomerizations of the cones in the retinal mosaic represent the first steps of visual processing.  To capture key properties of post-receptoral processing, contrast coding (REFS) and contrast normalization (REFS),
% Ask Johannes what he'd use as references for these broad principles
we processed the isomerations as follows:
First, we created a representation for each cone class on the full $51 \times 51$ pixel grid, we interpolated the responses of the cones in each class across the grid, using bilinear interpolation separately for each cone class.
% DOUBLE CHECK THAT THIS WAS DONE WITH SOMETHING LIKE INTERP2 WITH THE LINEAR OPTION.
This gives us a $51 \times 51$ cone response image for each cone class.
The three $51 \times 51$ sets of cone responses corresponding to the L, M and S cones were then reshaped into a single $7803 \times 1$ vector (7803 = 51 \times 51 \times 3). To model the contrast coding and contrast adaptation, we convert this vector to a contrast representation by subtracting out the mean and then dividing by it (contrast coding), followed by normalizing the vector to have unit length (contrast normalization).
The cone-response vectors corresponding to the N images for each case are concatenated column-wise to form a contrast normalized matrix $C_{7803\times N}$. The corresponding labels are represented by the row vector $Y_{1\times N}$.

@@

\subsubsection*{Linear Regression} For linear regression, we solve the equation $A_{1\times3}*C_{3\times N} = Y_{1\times N}$ for the regression parameters $A_{1\times3}$. Here, $C_{3\times N}$ is a matrix containing the contrast normalized L, M and S cone responses of the central patch of each image, extracted from $C_{7803\times N}$.  The columns corresponds to the N images in the dataset. The rows correspond to the contrast normalized L, M and S cones responses averaged over a 3$\times$3 pixel central region of the image containing the target object.  The (unknown) regression parameters $A$ are obtained using Matlab linear regression function. This model provides a baseline performance measure, which does not take the context in which the target image is seen into account.

\subsubsection*{SVD regression}
SVD regression is a generalization of principal component analysis that may be applied to labelled data \cite{ranganNotes}. Briefly, given an image dataset matrix $C_{M \times N}$ and labels $Y_{1\times N}$, the method finds unit vectors such that the projection of the data along these vectors maximizes the separation of the vectors with different labels, while clustering vectores with similar labels together. Mathematically, the method searches for a direction $\hat{u}$ that minimize the cost function:
\begin{align}
S = \sum\limits_{j,j'}\left(\hat{u}^\intercal C(:,j) - \hat{u}^\intercal C(:,j')\right)^2 \left(Y(j)-Y(j') \right)^2,
\end{align} 
where $C(:,j)$ represents the $j$'th column/image of the data matrix $C$. The direction of minimum cost is given by the eigenvector corresponding to the largest eigenvalue of the following matrix:
\begin{align}
N C Y_D^2 C^\intercal - 2(Y*\mathbbm{1})C Y_D C^\intercal + (Y*Y^\intercal) (CC^\intercal)-C Y_{D}^2\mathbbm{1}\mathbbm{1}^\intercal C^\intercal +2 C Y_{D} \mathbbm{1}\mathbbm{1}^\intercal Y_{D} C^\intercal - C \mathbbm{1}\mathbbm{1}^\intercal Y_{D}^2 C^\intercal.
\end{align}
Here, $Y_{D}$ is a diagonal matrix with the elements of the labels vector $Y$ along the diagonal, and $\mathbbm{1}$ is a $N\times 1$ vector with each element equal to one.
%% NEED TO DECIDE WHETHER WE ARE KEEPING THIS.  IF SO, NEED TO CLARIFY THAT WE CAN PROJECT ONTO MULTIPLE VECTORS AND THAT HAVING DONE SO, WE THEN DO BAYESIAN DECODING AS WITH AMA

% MAYBE ADD SOME INTRO TEXT THAT SAYS THAT FIRST ONE FINDS THE FILTERS AND THEN PROJECTS ONTO THEM, THEN DOES BAYESIAN DECODING OF THE RESPONSES.  BREAK INTO TWO SUBSECTIONS FOR CLARITY.
\subsubsection*{Accuracy maximization analysis}
AMA \cite{geisler2009optimal,burge2017accuracy} is a Bayesian ideal observer model for task specific dimensionality reduction. Given a labeled training set, a noisy response model, and a cost function, AMA returns a set of linear receptive fields that are optimal for performing a task. The task is implicitly defined in terms of the labels. The optimal filters that are returned by AMA are filters whose noisy response to the stimuli, as given by the specified response model, minimizes the cost function over the training set.
